\underline{Publications in refereed journals:}
\begin{enumerate}

\item{U. Kalmbach, T. Sch\"afer, T. S. Biro, U. Mosel,  Dirac sea effects in
the chiral quark soliton model, Nucl. Phys. A513 (1990) 621 }

\item{A. Hosaka, T. Sch\"afer, Spin structure of the nucleon in a
 nontopological chiral soliton model, Z. Phys. A337 (1990) 447}

\item{T. Sch\"afer, W. Weise, Neutral pion photoproduction at threshold and
explicit chiral symmetry breaking, Phys. Lett. B250 (1990) 6 }

\item{T. Sch\"afer, W. Weise, Threshold Pion Photoproduction and chiral
models of the nucleon, Nucl. Phys. A 531 (1991) 520, }

\item{T. Sch\"afer, V. Koch, G. E. Brown, Charge Symmetry Breaking
and the Neutron Proton Mass Difference, Nucl. Phys. A562 (1993) 644}

\item{T.~Sch\"afer, E.~V.~Shuryak, J.~J.~M.~Verbaarschot, Baryonic
Correlators in the Random Instanton Vacuum, Nucl.~Phys.~B412 (1994) 143}

\item{T.~Sch\"afer, E.~V.~Shuryak, Hadronic Wave Functions in the
Instanton Model, Phys.~Rev.~D50 (1994) 478}

\item{T.~Sch\"afer, E.~V.~Shuryak, J.~J.~M.~Verbaarschot, The Chiral
Phase Transition and Instanton-Antiinstanton Molecules, Phys.~Rev.~D51
(1995) 1267}

\item{T.~Sch\"afer, E.~V.~Shuryak, Instantons and Glueballs, 
Phys.~Rev.~Lett.~75 (1995) 1707}

\item{T.~Sch\"afer, E.~V.~Shuryak, Can hadrons survive the chiral phase
transition ?,  Phys.~Lett.~B356 (1995) 147}

\item{T.~Sch\"afer, E.~V.~Shuryak, The interacting instanton liquid in
QCD at zero and finite temperature, Phys. Rev. D53 (1996) 6522}

\item{T.~Sch\"afer, E.~V.~Shuryak, Hadronic correlators in the 
interacting instanton liquid, Phys. Rev. D54 (1996) 1099}

\item{T.~Sch\"afer, The $U(1)_A$ anomaly at finite temperature,
Phys. Lett. B389 (1996) 445}

\item{T.~Sch\"afer, Instantons and the Chiral Phase Transition
at non-zero Baryon Density, Phys. Rev. D, in press}

\newcounter{myitem}
\setcounter{myitem}{\value{enumi}}
\end{enumerate}


\underline{Review articles:}
\begin{enumerate}
\setcounter{enumi}{\value{myitem}}

\item{T.~Sch\"afer, E.~V.~Shuryak, The instanton liquid in QCD, 
Rev. Mod. Phys., in press}

\item{E.~V.~Shuryak, T.~Sch\"afer, The QCD vacuum as an instanton 
liquid, Ann. Rev. of Nucl. Part. Sc., in press}

\setcounter{myitem}{\value{enumi}}
\end{enumerate}


\underline{Preprints:}
\begin{enumerate}
\setcounter{enumi}{\value{myitem}}
 
\item{R. Rapp, T.~Sch\"afer, E. V. Shuryak, and M. Velkovsky,
Diquark Bose Condensates in high density matter and instantons, 
submitted to Phys. Rev. Lett.}

\setcounter{myitem}{\value{enumi}}
\end{enumerate}


\underline{Conference Proceedings:}
\begin{enumerate}
\setcounter{enumi}{\value{myitem}}

\item{T.~Sch\"afer, A Note on the Photoproduction Sigma Term, $\pi N$
news letter 3 (1991), 100}

\item{ T. Sch\"afer, Neutral Pion Photoproduction at Threshold and
Explicit Chiral Symmetry Breaking, Proceedings of the International
Workshop on Pions and Nuclei, E. Oset et al., eds., World Scientific 
(1991) }

\item{T.~Sch\"afer, Instantons and the Chiral Phase Transition, Proceedings 
of the 8th Meeting of the Division of Particles and Fields of the 
American Physical Society, S. Seidel, ed., World Scientific, Singapore,
1994}

\item{T.~Sch\"afer, The Chiral Phase Transition, Proceedings of 
Quark Matter 96, H. Specht et al., eds., Nucl. Phys. A610 (1996) 13c}

\item{A. Blotz, T.~Sch\"afer, E. Shuryak, Proton spin in the 
instanton liquid model, Proceedings of Baryons 95, B. F. Gibson
et al., eds., World Scientific, Singapore (1996)}

\item{E.~V.~Shuryak, T.~Sch\"afer, The Chiral Phase Transition in
the Instanton Liquid, Proceedings of Lattice 96, C. Bernard et al.,
eds., Nucl. Phys. B (Proc. Supp.) 53, 472 (1997) }

\item{T.~Sch\"afer, Progress in Instanton Liquid Calculations, Proceedings 
of the international conference on QCD: Chaos, Confinement and Collisions,
E. Fried and B. M\"uller, eds., World Scientific, Singapore (1997)}

\item{T.~Sch\"afer, QCD at finite baryon density: Instantons and Color
Superconductivity, Proceedings of Quark Matter 97, submitted to 
Nucl. Phys. A}

\end{enumerate}

\underline{Invited Talks:}
\begin{enumerate}

\item{ T. Sch\"afer, Neutral Pion Photoproduction at Threshold and
Explicit Chiral Symmetry Breaking, International Workshop on Pions 
and Nuclei, Peniscola, Spain (1991)}

\item{T.~Sch\"afer, A Note on the Photoproduction Sigma Term,
International Meeting on $\pi N$ Physics and the Structure of 
the Nucleon, Bad Honnef, Germany (1991)}

\item{T.~Sch\"afer, Instantons and the Chiral Phase Transition, 
Plenary Talk, APS march meeting, Washington (1995)}

\item{T.~Sch\"afer, The Chiral Phase Transition, Plenary Talk,
Quark Matter 96, Heidelberg, Germany (1996)}

\item{T.~Sch\"afer, Instantons and Color Superconductivity, 
Riken/BNL Workshop on Color Superconductivity, Instantons, 
and Parity Breaking at High Baryon Denisty, BNL (1997)}

\item{T.~Sch\"afer, QCD at finite baryon density, Parallel Talk,
Quark Matter 97, Tsukuba, Japan (1997)}

\end{enumerate}
