\documentstyle[12pt]{article}
\begin{document}   
\pagestyle{plain}

\centerline{\bf\Large Course Information PHY 506 - Fall 2006}
\vspace*{0.5cm}

 Physics 506 (Nuclear and Subatomic Physics) provides
an introduction to modern nuclear physics and is 
intended for advanced undergraduate or graduate 
students. A brief outline is provided below. 

\begin{enumerate}

\item{ {\bf Lecture:} MWF 12:25-1:15, rm204\\
 Instructor: T. Schaefer, Office BOM-210, Phone 513-7199\\
 email: Thomas\_Schaefer@ncsu.edu\\
 Office hours: We. 1:30-2:30 and by appointment.}

\item{ {\bf Homework:} Most mondays a set of homework problems 
 will be assigned. (You can also check the website.) The homework 
 is due monday the following week.}

\item{ {\bf Exams:} There will be one midterm exam, but no final.
Instead, we will assigns papers/projects and students will 
make presentations. }

\item{ {\bf Grade:} Your final grade will be determined by 
 weighting the various portions of the course as follows:\\ \\
 \begin{tabular}{ll}
 Midterm:       &  20\% \\
 Final Project :&  50\% \\
 Homework:      &  30\% \\
 \end{tabular} }

\item{ {\bf More Information:} Additional information related to 
 the course is available at 
{\tt http://wonka.physics.ncsu.edu/\~\ $\!\!\!$tmschaef/ }.
The official course syllabus can be found at
{\tt http://www.physics.ncsu.edu/
courses/tmschaef/Fall2006-506-001.html. }.}

\item{{\bf Textbook:} (not required, but recommended) Fundamentals 
in Nuclear Physics: From Nuclear Structure to Cosmology, by Jean-Louis 
Basdevant, James Rich, Michael Spiro (listed for \$68 at amazon.com). 
Supplemental references
 \begin{itemize}
 \item{K. Krane, Introductory Nuclear Physics}
 \item{S. S. M. Wong, Introductory Nuclear Physics}
 \item{B. Povh et al., Particles and Nuclei}
 \item{A. W. Thomas, W. Weise, The Structure of the Nucleon}
 \end{itemize}

There are many more advanced text books, for example A.~Bohr
and B.~Mottelson, Nuclear Structure (2 Vols), and H.~Feshbach,
Theoretical Nuclear Physics (2 Vols). Some supplemental material 
will be handed out in class or posted on the website.}
 
\item{{\bf Rough outline:}
 \begin{itemize}
 \item{\underline{Nucleons and Nuclear Forces}\\
Properties of Nucleons and Nuclei\\
Nucleon-Nucleon forces and the deuteron\\
Conservation laws, isospin.}

\item{\underline{Nuclear Models and Nuclear Stability}\\
The Fermi Gas and Liquid Drop Models\\
Radioactivity, Fusion and Fission\\
The Shell Model}

\item{\underline{Weak Interactions and Neutrinos}\\
Weak Interactions: $\beta$-deacy, $e$ capture, etc\\
Quarks and leptons\\
Neutrino Physics}

\item{\underline{Fission and Fusion, Nuclear Astrophysics }\\
Fission reactions, nuclear energy\\
Fusion recations\\
Stellar burning}

\item{\underline{The Quark Structure of Matter}\\
Deep Inelatic Scattering and the Structure of the Nucleon\\
Heavy Ion Collisions and the Quark Gluon Plasma}



 \end{itemize}}
\end{enumerate}
\end{document}   
 

 


  

